---
title: "LocExt_MonteCarlo"
author: "Manon Felix"
date: "2025-11-15"
output: html_document
---


# =========================================================
# R Code for Local Stationarity in the Extremes
# Author: Manon FELIX
# Date: [December 2025]
# =========================================================



# 1. Load Packages

```{r}
# tikzDevice: export plots to LaTeX
library(tikzDevice)

# Set timezone to California
options(tz = "CA")

# Data wrangling
library(tibble)
library(tidyverse)
library(dplyr)

# Matrix calculations
library(matrixcalc)

# Extreme value and stable distributions
library(evd)
library(stabledist)

# Plotting
library(ggplot2)
library(ggtext)
```




# 2. Functions

# 2.1 Second Derivative

```{r}
# Computes the Hessian (second derivative) of a multivariate function
# Reference: Rdistance package
#
# Args:
#   x    : numeric vector (point at which to evaluate)
#   FUN  : function to compute Hessian of
#   eps  : small perturbation for numerical derivative (default 1e-8)
#   ...  : additional arguments for FUN
#
# Returns:
#   Hessian matrix (d x d)

secondDeriv <- function(x, FUN, eps = 1e-8, ...) {
  d <- length(x)   # number of dimensions

  # Ensure eps vector has correct length
  if(d > length(eps)){
    eps <- rep(eps, ceiling(d / length(eps)))[1:d]
  } else if(length(eps) > d){
    eps <- eps[1:d]
  }

  FUN <- match.fun(FUN)
  hess <- matrix(0, nrow = d, ncol = d)
  h <- ifelse(x == 0, eps^0.25, (eps^0.25) * x)

  for(i in 1:d){
    ei <- rep(0, d)
    ei[i] <- 1

    # Diagonal elements
    hess[i,i] <- (-FUN(x + 2*h*ei, ...) + 
                   16*FUN(x + h*ei, ...) - 
                   30*FUN(x, ...) +
                   16*FUN(x - h*ei, ...) - 
                   FUN(x - 2*h*ei, ...)) / (12 * h[i]^2)

    # Off-diagonal elements
    if((i+1) <= d){
      for(j in (i+1):d){
        ej <- rep(0, d)
        ej[j] <- 1

        hess[i,j] <- (FUN(x + h*ei + h*ej, ...) - 
                        FUN(x + h*ei - h*ej, ...) -
                        FUN(x - h*ei + h*ej, ...) + 
                        FUN(x - h*ei - h*ej, ...)) / (4 * h[i] * h[j])

        # Symmetry
        hess[j,i] <- hess[i,j]
      }
    }
  }

  return(hess)
}
```




# 2.2 Kernel Function


```{r}
# Kernel function for smoothing or density estimation
#
# Args:
#   u0 : numeric, evaluation point
#   n  : numeric, number of observations
#   b  : numeric, bandwidth
#
# Returns:
#   Numeric vector of kernel weights

kern.func <- function(u0, n, b) {
  u <- (u0 - (1:n)/n) / b
  K <- ifelse(abs(u) < 1/2, 6 * (0.25 - abs(u)^2), 0)
  return(K)
}
```


# 3. Time-varying M4 process

Description: Defines localized estimators, psi functions, plots, normalized sums, and optimal bandwidth calculations.



```{r}
# -----------------------------
# 1. Localized Estimator
# -----------------------------
locest_am <- function(u, bn, x, prob, marg = 2){
  
  x <- as.matrix(x)
  n <- ifelse(is.null(dim(x)), length(x), dim(x)[1])
  mn <- 1/(1 - prob)
  
  # Custom CDF using kernel
  custom_cdf <- function(y, u, X) {
    n <- length(X)
    kernel <- kern.func(u, n, bn)
    sum(kernel * (X < y)) / sum(kernel)
  }

  # Inverse CDF
  invcdf <- function(p, X, u) {
    uniroot(function(x) custom_cdf(x, u = u, X) - p, range(X))$root
  }
  
  row_max <- apply(abs(x), 1, max)
  level <- invcdf(prob, row_max, u)
  
  kernel <- kern.func(u, n, bn)
  
  if (marg == 1 || marg == 2) {
    estim <- if (is.null(dim(x))) {
      ind <- x > level
      mn/(n*bn) * sum(ind * kernel)
    } else {
      binary_rows <- (x[,marg] > level)
      mn/(n * bn) * sum(kernel * as.numeric(binary_rows))
    }
  } else {
    estim <- if (is.null(dim(x))) {
      ind <- x > level
      mn/(n*bn) * sum(ind * kernel)
    } else {
      binary_rows <- (x[,1] > level) & (x[,2] > level)
      mn/(n * bn) * sum(kernel * as.numeric(binary_rows))
    }
  }
  
  return(estim)
}

# -----------------------------
# 2. M4(2) Process: Parameter Function
# -----------------------------
psi <- function(u, j, l, k){
  if (k == 1){
    val <- (l + 1) * 15 + 10 * (1 - 1.5 * u)
  } else {
    val <- (2 * (l + 1)) / (1 - abs(0.9 * sin((pi/2) * j - pi * u))^k)
  }
  return(val)
}

# -----------------------------
# 3. Plot of psi functions
# -----------------------------
s <- 70
us <- seq(0.1, 0.9, length.out = s)

par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))

# (j=1, l=0)
plot(us, psi(us, j=1, l=0, k=1),
     type="l", col="blue", lwd=2,
     xlim=c(0,1), ylim=c(0,40),
     ylab="psi(u)", xlab="u",
     main="j=1, l=0")
lines(us, psi(us, j=1, l=0, k=2), col="red", lwd=2, lty=2)
legend("topright", legend=c("k=1","k=2"), col=c("blue","red"),
       lty=c(1,2), lwd=2, bty="n", cex=0.8)

# (j=1, l=1)
plot(us, psi(us, j=1, l=1, k=1),
     type="l", col="blue", lwd=2,
     xlim=c(0,1), ylim=c(0,40),
     ylab="psi(u)", xlab="u",
     main="j=1, l=1")
lines(us, psi(us, j=1, l=1, k=2), col="red", lwd=2, lty=2)

# (j=0, l=0)
plot(us, psi(us, j=0, l=0, k=1),
     type="l", col="blue", lwd=2,
     xlim=c(0,1), ylim=c(0,40),
     ylab="psi(u)", xlab="u",
     main="j=0, l=0")
lines(us, psi(us, j=0, l=0, k=2), col="red", lwd=2, lty=2)

# (j=0, l=1)
plot(us, psi(us, j=0, l=1, k=1),
     type="l", col="blue", lwd=2,
     xlim=c(0,1), ylim=c(0,60),
     ylab="psi(u)", xlab="u",
     main="j=0, l=1")
lines(us, psi(us, j=0, l=1, k=2), col="red", lwd=2, lty=2)

# -----------------------------
# 4. Limiting Quantities
# -----------------------------
normalized_min_sum <- function(u, alpha, b) {
  numerator <- 0
  denominator <- 0

  for (l in 0:1) {
    for (j in 0:1) {
      psi1 <- psi(u, j, l, 1)
      psi2 <- psi(u, j, l, 2)
      min_val <- min(psi1 / b, psi2 / b)
      numerator <- numerator + min_val^alpha
      max_val <- max(psi1^alpha, psi2^alpha)
      denominator <- denominator + max_val
    }
  }
  return(numerator / denominator)
}

alpha <- 1.8
b <- 1
a <- 1
u_grid <- seq(0, 1, length.out = 200)
values <- sapply(u_grid, function(u) normalized_min_sum(u, alpha, b))

plot(u_grid, values, type = "l", col = "tomato", lwd = 2,
     xlab = "u", ylab = "Normalized Min Sum",
     main = expression(paste("Plot of Normalized ", min, " Sum vs ", u)))

seq_value <- function(u, alpha) {
  total <- 0
  for (l in 0:1) {
    for (j in 0:1) {
      max_k <- max(psi(u, j, l, 1)^alpha, psi(u, j, l, 2)^alpha)
      total <- total + max_k
    }
  }
  return(total)
}

# -----------------------------
# 5. Optimal Bandwidth
# -----------------------------
u_grid <- seq(0, 1, length.out = 500)
values <- sapply(u_grid, function(u) secondDeriv(u, normalized_min_sum, alpha = alpha, b = 1))

plot(u_grid, values, type = "l", col = "tomato", lwd = 2,
     xlab = "u", ylab = "Second Derivative",
     main = "Second derivative of normalized_min_sum")

n <- 9000
prob <- 0.98
mn <- 1/(1 - prob)

band_opt <- function(u){
  (mn/n)^(1/5) * (0.05 * (normalized_min_sum(u, alpha, b)+ 2) / 
                   (mn * secondDeriv(u, normalized_min_sum, alpha = alpha, b = 1) * 0.05)^2)^(1/5)
}

delta <- 0.2
u_grid <- seq(0 + delta, 1 - delta, length.out = 1000)
values <- sapply(u_grid, function(u) band_opt(u))
plot(u_grid, values, main = "Second derivative based bandwidth")

a1 <- function(u) {0.5 * mn * secondDeriv(u, normalized_min_sum, alpha = alpha, b = 1) * 0.05}
a2 <- function(u) {0.05 * (normalized_min_sum(u, alpha, b) + 2)}

bnopt <- function(u){
  (mn/n * a2(u) / (4 * a1(u)^2))^(1/5)
}

IMSE <- function(bn){
  u_grid <- seq(0 + delta, 1 - delta, length.out = 200)
  values <- sapply(u_grid, function(u) bn^4 * a1(u) + mn/(n*bn) * a2(u))
  pracma::trapz(u_grid, values)
}

bn_grid <- seq(0.03, 0.25, length.out = 100)
values <- sapply(bn_grid, function(bn) IMSE(bn))

pdf("/Users/manonfelix/Library/CloudStorage/OneDrive-Personnel/PhD/Research/Extrem_Manon/figures/bnIMSE.pdf", width = 8, height = 6)
plot(bn_grid, values, ylab = "IMSE", xlab = "bn", type = "l", main = "IMSE for different values of the bandwidth")
abline(v = bn_grid[which.min(values)], col = "red")
dev.off()

bnopt <- bn_grid[which.min(values)]
```


# 3.1 Functional Boxplots: Quantile Estimation


```{r}
# -----------------------------
# Settings and Parameters
# -----------------------------
prob <- 0.98
alpha <- 1.8
mn <- 1 / (1 - prob)
bn <- bnopt   # Optimal bandwidth from previous calculations
s <- 80       # Number of grid points for u
us <- seq(0.2, 0.8, length.out = s)
mt <- 1       # Number of Monte Carlo simulations
n <- 9000     # Sample size
set.seed(123)

# -----------------------------
# Define Quantile Functions
# -----------------------------
# Kernel-based CDF
custom_cdf <- function(y, u, X) {
  kernel <- kern.func(u, n, bn)
  binary_rows <- (X < y)
  1 / (n * bn) * sum(kernel * as.numeric(binary_rows))
}

# Inverse CDF using root finding
invcdf <- function(p, X, u = u0) {
  uniroot(function(x) custom_cdf(x, u = u, X) - p, range(X))$root
}

# Empirical (global) CDF
custom_cdf_stat <- function(y, X) {
  n <- length(X)
  binary_rows <- (X < y)
  sum(binary_rows) / n
}

# Inverse empirical CDF
invcdf_stat <- function(p, X) {
  uniroot(function(x) custom_cdf_stat(x, X) - p, range(X))$root
}

# -----------------------------
# Monte Carlo Simulation
# -----------------------------
estQlocal_mat <- matrix(NA, nrow = mt, ncol = s)
estQglobal_mat <- matrix(NA, nrow = mt, ncol = s)

for (b in 1:mt) {
  # Generate Frechet noise
  df <- as.data.frame(matrix(rfrechet(n * 4, shape = alpha), nrow = n, ncol = 4))
  
  # Random constants
  C_values <- sample(1:50, 4, replace = TRUE)
  
  # Simulate X_b
  X_b <- matrix(NA, nrow = n, ncol = 2)
  for (k in 1:2){
    for (i in 1:n){
      u <- i / n
      X_b[i, k] <- max(
        max((psi(u, j = 0, l = 0, k) + C_values[1]/n) * df[i, 1],
            (psi(u, j = 1, l = 0, k) + C_values[2]/n) * df[i, 2]),
        max((psi(u, j = 0, l = 1, k) + C_values[3]/n) * df[i, 3],
            (psi(u, j = 1, l = 1, k) + C_values[4]/n) * df[i, 4])
      )
    }
  }
  
  # Compute quantiles
  row_max_b <- apply(abs(X_b), 1, max)
  estQglobal_mat[b, ] <- invcdf_stat(p = prob, X = row_max_b)
  for (i in 1:s) {
    estQlocal_mat[b, i] <- invcdf(p = prob, X = row_max_b, u = us[i])
  }
  
  print(b)
}

# Save results
save(estQglobal_mat, estQlocal_mat, file = "/Users/manonfelix/Library/CloudStorage/OneDrive-Personnel/PhD/Research/Extrem_Manon/figures/estQ_matrices.RData")
load("/Users/manonfelix/Library/CloudStorage/OneDrive-Personnel/PhD/Research/Extrem_Manon/figures/estQ_matrices.RData")

# -----------------------------
# Convert to Long Data Frames
# -----------------------------
# TIME-VARYING
estQlocal_df <- as_tibble(estQlocal_mat)
colnames(estQlocal_df) <- paste0("u_", round(us, 4))
estQlocal_df <- estQlocal_df %>% 
  mutate(sim = 1:mt) %>%
  pivot_longer(-sim, names_to = "u", values_to = "estimate") %>%
  mutate(u = as.numeric(gsub("u_", "", u)))

# FIXED
estQglobal_df <- as_tibble(estQglobal_mat)
colnames(estQglobal_df) <- paste0("u_", round(us, 3))
estQglobal_df <- estQglobal_df %>% 
  mutate(sim = 1:mt) %>%
  pivot_longer(-sim, names_to = "u", values_to = "estimate") %>%
  mutate(u = as.numeric(gsub("u_", "", u)))

# -----------------------------
# True Sequence for Comparison
# -----------------------------
sigma <- sapply(us, function(u) seq_value(u, alpha))
true_seq <- (1/(1 - prob) * sigma)^(1/alpha)
comparison_df <- data.frame(u = us, value = true_seq)

# -----------------------------
# Compute Summary Statistics
# -----------------------------
# TIME-VARYING
summary_stats <- estQlocal_df %>%
  group_by(u) %>%
  summarise(
    median = median(estimate, na.rm = TRUE),
    q25 = quantile(estimate, 0.25, na.rm = TRUE),
    q75 = quantile(estimate, 0.75, na.rm = TRUE)
  )

# FIXED
summary_fixed <- estQglobal_df %>%
  group_by(u) %>%
  summarise(
    median = median(estimate, na.rm = TRUE),
    q25 = quantile(estimate, 0.25, na.rm = TRUE),
    q75 = quantile(estimate, 0.75, na.rm = TRUE)
  )

# -----------------------------
# Plot Functional Boxplots
# -----------------------------
col_fixed <- "firebrick3"
col_local <- "steelblue3"

estQ_plot <- ggplot() +
  # Fixed estimator
  geom_ribbon(data = summary_fixed, aes(x = u, ymin = q25, ymax = q75), fill = col_fixed, alpha = 0.2) +
  geom_line(data = summary_fixed, aes(x = u, y = median), color = col_fixed, size = 1.2) +
  geom_line(data = summary_fixed, aes(x = u, y = q25), color = col_fixed, linetype = "dashed", size = 0.8) +
  geom_line(data = summary_fixed, aes(x = u, y = q75), color = col_fixed, linetype = "dashed", size = 0.8) +
  
  # Local estimator
  geom_ribbon(data = summary_stats, aes(x = u, ymin = q25, ymax = q75), fill = col_local, alpha = 0.2) +
  geom_line(data = summary_stats, aes(x = u, y = median), color = col_local, size = 1.2) +
  geom_line(data = summary_stats, aes(x = u, y = q25), color = col_local, linetype = "dashed", size = 0.8) +
  geom_line(data = summary_stats, aes(x = u, y = q75), color = col_local, linetype = "dashed", size = 0.8) +
  
  # Comparison line
  geom_line(data = comparison_df, aes(x = u, y = value), linetype = "dotted", color = "black", size = 0.8) +
  
  # Labels & theme
  labs(title = "Estimation of",
       subtitle = expression("time-varying quantile " * q(u, 1 - 1/m[n])),
       x = "u",
       y = "Estimated value") +
  theme_minimal(base_size = 14, base_family = "Times") +
  theme(legend.position = "none",
        plot.title = element_text(size = 16, family = "Times"),
        axis.title = element_text(size = 14, family = "Times"),
        axis.text = element_text(size = 14, family = "Times"))

estQ_plot

```


# 3.2 Functional Boxplot: Joint Measure ν1(u, C)

```{r}
library(tidyverse)
library(boot)

# -----------------------------
# Settings and Parameters
# -----------------------------
prob <- 0.98
alpha <- 1.8
mn <- 1 / (1 - prob)
bn <- bnopt       # Optimal bandwidth from previous calculations
s <- 80           # Number of grid points for u
us <- seq(0.2, 0.8, length.out = s)
B <- 1000         # Number of Monte Carlo simulations
n <- 9000         # Sample size
set.seed(123)

# -----------------------------
# Define Estimator Functions
# -----------------------------
# Direct joint estimator function for global measure
est.fun <- function(tsb) {
  mn <- 1 / (1 - prob)
  X_mat <- as.matrix(tsb)
  row_max <- apply(abs(X_mat), 1, max)
  level <- quantile(row_max, prob)
  mn * mean((X_mat[,1] > level) & (X_mat[,2] > level))
}

# -----------------------------
# Monte Carlo Simulation
# -----------------------------
estPglobal_mat <- matrix(NA, nrow = B, ncol = s)
estPlocal_mat <- matrix(NA, nrow = B, ncol = s)

for (b in 1:B) {
  # Generate Frechet noise
  df <- as.data.frame(matrix(rfrechet(n * 4, shape = alpha), nrow = n, ncol = 4))
  
  # Random constants
  C_values <- sample(1:50, 4, replace = TRUE)
  
  # Simulate X_b
  X_b <- matrix(NA, nrow = n, ncol = 2)
  for (k in 1:2) {
    for (i in 1:n) {
      u <- i / n
      X_b[i, k] <- max(
        max((psi(u, j = 0, l = 0, k) + C_values[1]/n) * df[i, 1],
            (psi(u, j = 1, l = 0, k) + C_values[2]/n) * df[i, 2]),
        max((psi(u, j = 0, l = 1, k) + C_values[3]/n) * df[i, 3],
            (psi(u, j = 1, l = 1, k) + C_values[4]/n) * df[i, 4])
      )
    }
  }
  
  # Compute estimators over grid
  estPglobal_mat[b, ] <- est.fun(X_b)
  for (i in 1:s) {
    estPlocal_mat[b, i] <- locest_am(us[i], bn = bn, x = X_b, prob = prob, marg = 0)
  }
}

# Save/load results
save(estPglobal_mat, estPlocal_mat, file = "/Users/manonfelix/Library/CloudStorage/OneDrive-Personnel/PhD/Research/Extrem_Manon/figures/estP_matrices.RData")
load("/Users/manonfelix/Library/CloudStorage/OneDrive-Personnel/PhD/Research/Extrem_Manon/figures/estP_matrices.RData")

# -----------------------------
# Convert to Long Data Frames
# -----------------------------
# Time-varying estimator
estPlocal_df <- as_tibble(estPlocal_mat)
colnames(estPlocal_df) <- paste0("u_", round(us, 3))
estPlocal_df <- estPlocal_df %>% 
  mutate(sim = 1:B) %>%
  pivot_longer(-sim, names_to = "u", values_to = "estimate") %>%
  mutate(u = as.numeric(gsub("u_", "", u)))

# Fixed/global estimator
estPglobal_df <- as_tibble(estPglobal_mat)
colnames(estPglobal_df) <- paste0("u_", round(us, 3))
estPglobal_df <- estPglobal_df %>% 
  mutate(sim = 1:B) %>%
  pivot_longer(-sim, names_to = "u", values_to = "estimate") %>%
  mutate(u = as.numeric(gsub("u_", "", u)))

# -----------------------------
# Optional Comparison Line
# -----------------------------
comparison_vals <- sapply(us, function(u) normalized_min_sum(u, alpha, b))
comparison_df <- data.frame(u = us, value = comparison_vals)

# -----------------------------
# Compute Summary Statistics
# -----------------------------
# Time-varying estimator
summary_stats <- estPlocal_df %>%
  group_by(u) %>%
  summarise(
    median = median(estimate, na.rm = TRUE),
    q25 = quantile(estimate, 0.25, na.rm = TRUE),
    q75 = quantile(estimate, 0.75, na.rm = TRUE)
  )

# Fixed/global estimator
summary_fixed <- estPglobal_df %>%
  group_by(u) %>%
  summarise(
    median = median(estimate, na.rm = TRUE),
    q25 = quantile(estimate, 0.25, na.rm = TRUE),
    q75 = quantile(estimate, 0.75, na.rm = TRUE)
  )

# -----------------------------
# Plot Functional Boxplots
# -----------------------------
col_fixed <- "firebrick3"    # Fixed/global estimator
col_local <- "steelblue3"    # Time-varying/local estimator

estP_plot <- ggplot() +
  # Fixed estimator
  geom_ribbon(data = summary_fixed, aes(x = u, ymin = q25, ymax = q75), fill = col_fixed, alpha = 0.2) +
  geom_line(data = summary_fixed, aes(x = u, y = median), color = col_fixed, size = 1.2) +
  geom_line(data = summary_fixed, aes(x = u, y = q25), color = col_fixed, linetype = "dashed", size = 0.8) +
  geom_line(data = summary_fixed, aes(x = u, y = q75), color = col_fixed, linetype = "dashed", size = 0.8) +
  
  # Local estimator
  geom_ribbon(data = summary_stats, aes(x = u, ymin = q25, ymax = q75), fill = col_local, alpha = 0.2) +
  geom_line(data = summary_stats, aes(x = u, y = median), color = col_local, size = 1.2) +
  geom_line(data = summary_stats, aes(x = u, y = q25), color = col_local, linetype = "dashed", size = 0.8) +
  geom_line(data = summary_stats, aes(x = u, y = q75), color = col_local, linetype = "dashed", size = 0.8) +
  
  # Comparison line
  geom_line(data = comparison_df, aes(x = u, y = value), linetype = "dotted", color = "black", size = 0.8) +
  
  # Labels and theme
  labs(title = "Estimation of",
       subtitle = expression("time-varying tail measure " * nu[1](u, C)),
       x = "u",
       y = "Estimated value") +
  theme_minimal(base_size = 14, base_family = "Times") +
  theme(legend.position = "none",
        plot.title = element_text(size = 16, family = "Times"),
        axis.title = element_text(size = 14, family = "Times"),
        axis.text = element_text(size = 14, family = "Times"))

estP_plot


```


# 4. Time-varying AR(1) process

```{r}
# =========================================================
# 4. Time-varying AR(1) Process + Quantile + Extremogram
# =========================================================

library(tidyverse)
library(stabledist)
library(gridExtra)

# -----------------------------
# AR(1) Parameter Function
# -----------------------------
s <- 50
us <- seq(0.2, 0.8, length.out = s)

# Sine-based time-varying coefficient
fct_sine <- function(t, T) {
  0.2 + 0.4*t/T + 0.2*(t/T)^2 
}

T <- length(us)
t <- 1:T
plot(us, fct_sine(t, T))

fct_sine_u <- function(u) {
  0.2 + 0.4*u + 0.2*(u)^2 
}

# Function for series sum (used for theoretical quantiles)
series_sum <- function(u, alpha) {
  r <- abs(fct_sine_u(u))^alpha
  1 / (1 - r)
}
```



# 4.1 Functional Boxplots: Extreme Quantile

```{r}
dim_para <- 1
alpha <- 1
prob <- 0.98
s <- 80
us <- seq(0.2, 0.8, length.out = s)
B <- 1000
n <- 9000
bn <- bnopt  # candidate bandwidth

# Custom CDF function
custom_cdf <- function(y, u, X) {
  kernel <- kern.func(u, n, bn)
  binary_rows <- (X < y)
  1 / (n * bn) * sum(kernel * as.numeric(binary_rows))
}

# Inverse CDF using root finding
invcdf <- function(p, X, u = u0) {
  uniroot(function(x) custom_cdf(x, u = u, X) - p, range(X))$root
}

# Empirical CDF (no kernel)
custom_cdf_stat <- function(y, X) {
  n <- length(X)
  binary_rows <- (X < y)
  sum(binary_rows) / n
}

invcdf_stat <- function(p, X) {
  uniroot(function(x) custom_cdf_stat(x, X) - p, range(X))$root
}

# Monte Carlo simulation for quantiles
estQlocal_mat <- matrix(NA, nrow = B, ncol = s)
estQglobal_mat <- matrix(NA, nrow = B, ncol = s)
set.seed(123)

for (b in 1:B) {
  print(b)
  eps <- rstable(n, alpha = alpha, beta = 0)
  X_b <- array(NA, dim = n)
  X_b[1:dim_para] <- 0 
  
  for (j in dim_para:(n-1)) {
    X_b[j + 1] <- rep(as.numeric(fct_sine(j+1, n)), dim_para) %*% X_b[(j-(dim_para-1)):j] + eps[j]
  }
  
  estQglobal_mat[b, ] <- invcdf_stat(p = prob, X = abs(X_b))
  for (i in 1:s) {
    estQlocal_mat[b, i] <- invcdf(p = prob, X = abs(X_b), u = us[i])
  }
}

save(estQglobal_mat, estQlocal_mat, file = "/Users/manonfelix/Library/CloudStorage/OneDrive-Personnel/PhD/Research/Extrem_Manon/figures/estQAR_matrices.RData")
load("/Users/manonfelix/Library/CloudStorage/OneDrive-Personnel/PhD/Research/Extrem_Manon/figures/estQAR_matrices.RData")

# Convert to long data frames
estQlocal_df <- as_tibble(estQlocal_mat) %>%
  set_names(paste0("u_", round(us, 3))) %>%
  mutate(sim = 1:B) %>%
  pivot_longer(-sim, names_to = "u", values_to = "estimate") %>%
  mutate(u = as.numeric(gsub("u_", "", u)))

estQglobal_df <- as_tibble(estQglobal_mat) %>%
  set_names(paste0("u_", round(us, 3))) %>%
  mutate(sim = 1:B) %>%
  pivot_longer(-sim, names_to = "u", values_to = "estimate") %>%
  mutate(u = as.numeric(gsub("u_", "", u)))

# Comparison line: theoretical quantile
comparison_df <- data.frame(u = us) %>%
  rowwise() %>%
  mutate(
    gamma = (series_sum(u, alpha))^(1/alpha),
    value = qstable((1 + prob)/2, alpha = alpha, beta = 0, gamma = gamma, delta = 0)
  ) %>%
  ungroup()

# Compute summary statistics for functional boxplots
summary_stats <- estQlocal_df %>%
  group_by(u) %>%
  summarise(median = median(estimate, na.rm = TRUE),
            q25 = quantile(estimate, 0.25, na.rm = TRUE),
            q75 = quantile(estimate, 0.75, na.rm = TRUE))

summary_fixed <- estQglobal_df %>%
  group_by(u) %>%
  summarise(median = median(estimate, na.rm = TRUE),
            q25 = quantile(estimate, 0.25, na.rm = TRUE),
            q75 = quantile(estimate, 0.75, na.rm = TRUE))

# Colors
col_fixed <- "firebrick3"
col_local <- "steelblue3"

# Plot functional boxplots
estQ_plotAR <- ggplot() +
  geom_ribbon(data = summary_fixed, aes(x = u, ymin = q25, ymax = q75), fill = col_fixed, alpha = 0.2) +
  geom_line(data = summary_fixed, aes(x = u, y = median), color = col_fixed, size = 1.2) +
  geom_line(data = summary_fixed, aes(x = u, y = q25), color = col_fixed, linetype = "dashed", size = 0.8) +
  geom_line(data = summary_fixed, aes(x = u, y = q75), color = col_fixed, linetype = "dashed", size = 0.8) +
  geom_ribbon(data = summary_stats, aes(x = u, ymin = q25, ymax = q75), fill = col_local, alpha = 0.2) +
  geom_line(data = summary_stats, aes(x = u, y = median), color = col_local, size = 1.2) +
  geom_line(data = summary_stats, aes(x = u, y = q25), color = col_local, linetype = "dashed", size = 0.8) +
  geom_line(data = summary_stats, aes(x = u, y = q75), color = col_local, linetype = "dashed", size = 0.8) +
  geom_line(data = comparison_df, aes(x = u, y = value), linetype = "dotted", color = "black", size = 0.8) +
  labs(title = "Estimation of",
       subtitle = expression("time-varying quantile " * q(u, 1 - 1/m[n])),
       x = "u", y = "Estimated value") +
  theme_minimal(base_size = 14, base_family = "Times") +
  theme(legend.position = "none",
        plot.title = element_text(size = 16, family = "Times"),
        axis.title = element_text(size = 14, family = "Times"),
        axis.text = element_text(size = 14, family = "Times"))

estQ_plotAR
```



# 4.2 Functional Boxplots: Extremogram

```{r}
est.fun.stat <- function(tsb, h = 0) {
  mn <- 1 / (1 - prob)
  x <- as.matrix(tsb)
  n <- nrow(x)
  row_max <- apply(abs(x), 1, max)
  level <- quantile(row_max, probs = prob)
  if (h >= 0) {
    X1 <- x[1:(n-h)]
    X2 <- x[(1+h):n]
  } else {
    hh <- abs(h)
    X1 <- x[(1-h):n]
    X2 <- x[1:(n+h)]
  }
  band_condition <- (X1 > level) & (X2 > level)
  mn * mean(band_condition)
}

# Monte Carlo simulation for extremogram
estElocal_mat <- matrix(NA, nrow = B, ncol = s)
estEglobal_mat <- matrix(NA, nrow = B, ncol = s)
h <- 1

for (b in 1:B) {
  eps <- rstable(n, alpha = alpha, beta = 0)
  X_b <- array(NA, dim = n)
  X_b[1:dim_para] <- 0 
  for (j in dim_para:(n-1)) {
    X_b[j + 1] <- rep(as.numeric(fct_sine(j+1, n)), dim_para) %*% X_b[(j-(dim_para-1)):j] + eps[j]
  }
  estEglobal_mat[b, ] <- est.fun.stat(X_b, h = h)
  
  leveltv <- rep(NA, s)
  for (i in 1:s) {
    x <- as.matrix(X_b)
    n <- nrow(x)
    row_max <- apply(abs(x), 1, max)
    leveltv[i] <- invcdf(prob, row_max, u = us[i], bn = bn)
    
    if (h >= 0) {
      X1 <- x[1:(n-h)]
      X2 <- x[(1+h):n]
      kern <- kern.func(u0 = us[i], n = n-h, b = bn)
    } else {
      X1 <- x[(1-h):n]
      X2 <- x[1:(n+h)]
      kern <- kern.func(u0 = us[i], n = n+h, b = bn)
    }
    
    band_condition <- (X1 > leveltv[i]) & (X2 > leveltv[i])
    estElocal_mat[b, i] <- mn / (length(X1) * bn) * sum(kern * band_condition)
  }
  print(b)
}

save(estEglobal_mat, estElocal_mat, file = "/Users/manonfelix/Library/CloudStorage/OneDrive-Personnel/PhD/Research/Extrem_Manon/figures/estE_matrices.RData")
load("/Users/manonfelix/Library/CloudStorage/OneDrive-Personnel/PhD/Research/Extrem_Manon/figures/estE_matrices.RData")

# Convert to long data frames
estElocal_df <- as_tibble(estElocal_mat) %>%
  set_names(paste0("u_", round(us, 3))) %>%
  mutate(sim = 1:B) %>%
  pivot_longer(-sim, names_to = "u", values_to = "estimate") %>%
  mutate(u = as.numeric(gsub("u_", "", u)))

estEglobal_df <- as_tibble(estEglobal_mat) %>%
  set_names(paste0("u_", round(us, 3))) %>%
  mutate(sim = 1:B) %>%
  pivot_longer(-sim, names_to = "u", values_to = "estimate") %>%
  mutate(u = as.numeric(gsub("u_", "", u)))

comparison_df <- data.frame(u = us) %>%
  rowwise() %>%
  mutate(value = (fct_sine_u(u))^(alpha*h)/2) %>%
  ungroup()

summary_stats <- estElocal_df %>%
  group_by(u) %>%
  summarise(median = median(estimate, na.rm = TRUE),
            q25 = quantile(estimate, 0.25, na.rm = TRUE),
            q75 = quantile(estimate, 0.75, na.rm = TRUE))

summary_fixed <- estEglobal_df %>%
  group_by(u) %>%
  summarise(median = median(estimate, na.rm = TRUE),
            q25 = quantile(estimate, 0.25, na.rm = TRUE),
            q75 = quantile(estimate, 0.75, na.rm = TRUE))

col_fixed <- "firebrick3"
col_local <- "steelblue3"

estE_plotAR <- ggplot() +
  geom_ribbon(data = summary_fixed, aes(x = u, ymin = q25, ymax = q75), fill = col_fixed, alpha = 0.2) +
  geom_line(data = summary_fixed, aes(x = u, y = median), color = col_fixed, size = 1.2) +
  geom_line(data = summary_fixed, aes(x = u, y = q25), color = col_fixed, linetype = "dashed", size = 0.8) +
  geom_line(data = summary_fixed, aes(x = u, y = q75), color = col_fixed, linetype = "dashed", size = 0.8) +
  geom_ribbon(data = summary_stats, aes(x = u, ymin = q25, ymax = q75), fill = col_local, alpha = 0.2) +
  geom_line(data = summary_stats, aes(x = u, y = median), color = col_local, size = 1.2) +
  geom_line(data = summary_stats, aes(x = u, y = q25), color = col_local, linetype = "dashed", size = 0.8) +
  geom_line(data = summary_stats, aes(x = u, y = q75), color = col_local, linetype = "dashed", size = 0.8) +
  geom_line(data = comparison_df, aes(x = u, y = value), linetype = "dotted", color = "black", size = 0.8) +
  labs(title = "Estimation of",
       subtitle = expression("time-varying extremogram " * nu[h+1](u, A,B)),
       x = "u", y = "Estimated value") +
  theme_minimal(base_size = 14, base_family = "Times") +
  theme(legend.position = "none",
        plot.title = element_text(size = 16, family = "Times"),
        axis.title = element_text(size = 14, family = "Times"),
        axis.text = element_text(size = 14, family = "Times"))

estE_plotAR



```



